---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# actuary

<!-- badges: start -->
[![R-CMD-check](https://github.com/andrewwilson201/actuary/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/andrewwilson201/actuary/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `actuary` package provides actuarial modeling tools for Monte Carlo loss simulations, loss reserving, and reinsurance layer loss calculations. It enables users to generate stochastic loss datasets with customisable frequency and severity distributions, fit development patterns to claim triangles, and calculate reinsurance losses for occurrence and aggregate layers with user-defined retentions, limits, and reinstatements. For development pattern selection, the package includes a machine learning approach that evaluates multiple reserving models using holdout validation to identify the best-fitting pattern based on predictive accuracy.


## Installation

You can install the development version of actuary from [GitHub](https://github.com/andrewwilson201/actuary) with:

``` r
# install.packages("pak")
pak::pak("andrewwilson201/actuary")
```

## Examples

### Monte Carlo Loss Simulations
Generate simulated loss datasets using custom frequency and severity distributions.

```{r monte_carlo}
library(actuary)

# Simulate 100,000 years of loss data with Poisson frequency (mean = 5) and Lognormal severity (meanlog = 13, sdlog = 0.5)
sim_data <- create_simulations(
  num_sims = 100000, 
  mean_freq = 5, 
  sev_dist = "lognorm2", 
  sev_param1 = 13, 
  sev_param2 = 0.5
)

head(sim_data)

```

### Loss Reserving with Development Patterns
Fit a development pattern to a cumulative claims triangle. 

```{r loss_reserving}
# Example dataset
triangle_data <- actuary::triangle_data

# Fit chain-ladder model with optional smoothing
fit <- fit_development_pattern(
  cohort_var = uw_year, 
  dev_var = dev_year, 
  weighting_var = claim_number, 
  data = triangle_data, 
  smooth_from = 4
)

# View fitted development pattern
plot_development_pattern(fit)
```

### Machine Learning for Development Pattern Selection
Automatically select the best-fitting development pattern based on holdout validation.

```{r mlr}
# Define multiple reserving model parameters for evaluation
params <- expand.grid(
  exclude_last_diag = c(TRUE, FALSE),
  smooth_from = c(3, 4),
  exclude_high = c(TRUE),
  exclude_low = c(TRUE),
  selected_curve = c("weibull", "exponential_decay"),
  num_periods = c(3, 4),
  future_dev_periods = c(0, 12)
)

# Run machine learning model selection
ml_result <- fit_development_pattern_ml(
  cohort_var = uw_year, 
  dev_var = dev_year, 
  weighting_var = claim_number, 
  data = triangle_data, 
  dev_period_length = 1, 
  dev_period_units = 12, 
  params_data = params
)

# View ranked results
ml_result$results

# Plot best-fitting development pattern
plot_development_pattern(ml_result$best_fit)
```


### Reinsurance Layer Loss Calculations
Calculate occurrence and aggregate reinsurance losses.

```{r layer_losses}
# Example loss dataset
losses <- data.frame(
  year = rep(1:5, each = 5),
  amount = c(1e6, 3e6, 5e6, 7e6, 9e6, 2e6, 4e6, 6e6, 8e6, 10e6, 
             1.5e6, 3.5e6, 5.5e6, 7.5e6, 9.5e6, 2.5e6, 4.5e6, 6.5e6, 8.5e6, 10.5e6,
             3e6, 6e6, 9e6, 12e6, 15e6)
)

# Calculate 10M x 10M occurrence loss with one reinstatement
losses$occurrence_layer <- layer_loss(
  year = losses$year, 
  loss = losses$amount, 
  retention = 10e6, 
  limit = 10e6, 
  reinstatements = 1, 
  type = "occurrence"
)

# Calculate 10M x 10M aggregate layer loss with no reinstatements
losses$aggregate_layer <- layer_loss(
  year = losses$year, 
  loss = losses$amount, 
  retention = 10e6, 
  limit = 10e6, 
  reinstatements = 0, 
  type = "aggregate"
)

head(losses)

```


